Testing Forest Area as a random effect

### 2022-10-03 paper/Forest_and_Water/Test_model_totalForestArea.tex
resolve comment about percent total forest

\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Supplementary Information part 3: Testing different forest cover relationships},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{Supplementary Information part 3: Testing different forest cover relationships}
    \author[]{R. Willem Vervoort%
  %
  \fnref{1}}
   \ead{willem.vervoort@sydney.edu.au} 
    \author[]{Eliana Nervi}
   \ead{eliananervif@gmail.com} 
    \author[]{Jimena Alonso}
   \ead{jalonso@fing.edu.uy} 
      \cortext[cor1]{Corresponding author}
    \fntext[1]{Corresponding Author}

@@

In this document we tested the following, which is outlined in the methods of the paper.
The changes in forest cover contain both positive (forestation) and negative values (deforestation). In \citet{zhang2017}, these changes were jointly analysed, assuming the effect on the change in flow was linear and the effect of removing forest cover was the same as an equivalent addition of forest cover. Here we test two variation of this approach, leading to three approaches:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Joint analysis and assuming a linear relationship between the change in streamflow and the change in forest cover, following \citet{zhang2017}.
\item
  The impact of an increase in forest cover can be different from the same fractional decrease in forest cover. The question becomes how best to analyse this. One approach would be to allow a different slope and a different intercept for the decreases relative to the increases.
\item
  A second approach is to test the change in forest cover as a non-linear relationship in the Generalised additive model (GAM).
\end{enumerate}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{approach-1}{%
\subsection{Approach 1}\label{approach-1}}

To estimate how the change in streamflow is affected by the change in forest cover, while considering the effects of the other variables, we applied generalised additive modelling (GAM) \citep{wood2006}.

The general model tested is:

\begin{align}
\Delta Qf \% \sim &~ \Delta \% forest~cover + \notag \\ 
& \sum{X_i} + \sum{s(Z_i)} + \varepsilon \label{eq:eq1}
\end{align}

\hypertarget{approach-2}{%
\subsection{Approach 2}\label{approach-2}}

Deforestation is different from reforestation. This can be tested by converting all the change in forest cover data to positive values, and an additional binary column (\(sign_{forest cover}\)) can be included indicating whether it was a forest cover increase or decrease. In the model, the parameter for \(sign_{forest cover}\) will indicate the difference in the changes in flow for increases in forest cover compared to decreases in forest cover. The disadvantage of this approach is that the relationship with forest cover becomes discontinuous at the origin (0 change in forest cover).

This results in the following model:

\begin{align}
\Delta Qf \% \sim &~ \Delta \% forest~cover_{positive} + sign_{forest cover} + \notag \\ 
& \sum{X_i} + \sum{s(Z_i)} + \varepsilon \label{eq:eq2}
\end{align}

\hypertarget{approach-3}{%
\subsection{Approach 3}\label{approach-3}}

The relationship between the change in forest cover and change in streamflow is non-linear. The non-linear approach involves adding a smoothing function the the change in forestry. A shrinkage spline \citep{wood2006} is applied to allow the non-linear effect to be removed. Because a shrinkage penalty is used, this will also test the non-linear assumption and allows the variable for forest cover to be continuous. The disadvantage of this approach is that the relationship between forest cover and change in flow is less easy to interpret, as the non-linear fit in the GAM has no direct parametric form. This leads to the following model:

\begin{align}
\Delta Qf \% \sim &~ s(\Delta \% forest~cover) + \notag \\ 
& \sum{X_i} + \sum{s(Z_i)} + \varepsilon \label{eq:eq3}
\end{align}

\hypertarget{preliminary-data-management}{%
\subsection{Preliminary data management}\label{preliminary-data-management}}

First we will read in the data

We will combine the different tables, but will keep an indicator to see where the data are from.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Zhang\_small}\SpecialCharTok{$}\NormalTok{From }\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(Zhang\_small}\SpecialCharTok{$}\NormalTok{From)}
\NormalTok{Zhang\_small}\SpecialCharTok{$}\NormalTok{To }\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(Zhang\_small}\SpecialCharTok{$}\NormalTok{To)}
\NormalTok{Zhang\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(Zhang\_large,Zhang\_small) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dataset =} \StringTok{"original Zhang et al data"}\NormalTok{)}
\NormalTok{new\_data }\OtherTok{\textless{}{-}}\NormalTok{ new\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dataset =} \StringTok{"new data"}\NormalTok{)}
\NormalTok{All\_data }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(Zhang\_all, new\_data)}
\end{Highlighting}
\end{Shaded}

\hypertarget{implementing-the-changes-to-the-overall-data}{%
\subsection{Implementing the changes to the overall data}\label{implementing-the-changes-to-the-overall-data}}

The following code implements the changes described in the Supplementary data part 1. However, many of the changes were implemented manually into the data set. These are simply the remaining changes not implemented manually.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  removing the duplicates.
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{All\_data }\OtherTok{\textless{}{-}}\NormalTok{ All\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Possible duplicate}\StringTok{\textasciigrave{}} \OtherTok{=} 
           \FunctionTok{ifelse}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Possible duplicate}\StringTok{\textasciigrave{}}\NormalTok{)}\SpecialCharTok{==}\NormalTok{T,}\DecValTok{0}\NormalTok{,}\StringTok{\textasciigrave{}}\AttributeTok{Possible duplicate}\StringTok{\textasciigrave{}}\NormalTok{),}
         \StringTok{\textasciigrave{}}\AttributeTok{Possible duplicate}\StringTok{\textasciigrave{}} \OtherTok{=} \FunctionTok{as.numeric}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Possible duplicate}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Possible duplicate}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  calculating the dryness
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# calculate dryness index}
\NormalTok{All\_data }\OtherTok{\textless{}{-}}\NormalTok{ All\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Dryness =}\NormalTok{ E0}\SpecialCharTok{/}\NormalTok{Pa\_mm)}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  remove watershed 1 (the Amazon) from the analysis
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{All\_data }\OtherTok{\textless{}{-}}\NormalTok{ All\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Watershed \#}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  remove data set 188 and 254 Kamakia and Sambret
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{All\_data }\OtherTok{\textless{}{-}}\NormalTok{ All\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Watershed \#}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \DecValTok{188}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Watershed \#}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \DecValTok{254}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  add a column that indicates forst loss of forest gain
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{All\_data }\OtherTok{\textless{}{-}}\NormalTok{ All\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{forest\_sign =} \FunctionTok{ifelse}\NormalTok{(DeltaF\_perc }\SpecialCharTok{\textless{}} \DecValTok{0}\NormalTok{, }\StringTok{"Forest Cover Loss"}\NormalTok{,}
                              \StringTok{"Forest Cover Gain"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\hypertarget{approach-1-linear-relationship-across-all-changes-in-forest-cover}{%
\subsection{Approach 1: linear relationship across all changes in forest cover}\label{approach-1-linear-relationship-across-all-changes-in-forest-cover}}

First the variable \texttt{length} is added, representing the number of years that the data represents.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{All\_data2 }\OtherTok{\textless{}{-}}\NormalTok{ All\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Forest\_Sign =} \FunctionTok{ifelse}\NormalTok{(DeltaF\_perc }\SpecialCharTok{\textless{}} \DecValTok{0}\NormalTok{,}
                              \StringTok{"Decrease"}\NormalTok{, }\StringTok{"Increase"}\NormalTok{),}
         \AttributeTok{DeltaF\_perc\_pos =} \FunctionTok{ifelse}\NormalTok{(DeltaF\_perc }\SpecialCharTok{\textless{}} \DecValTok{0}\NormalTok{,}
                                  \SpecialCharTok{{-}}\DecValTok{1}\SpecialCharTok{*}\NormalTok{DeltaF\_perc,}
\NormalTok{                                  DeltaF\_perc))}
\NormalTok{All\_data2 }\OtherTok{\textless{}{-}}\NormalTok{ All\_data2 }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{length =}\NormalTok{ To }\SpecialCharTok{{-}}\NormalTok{ From,}
         \AttributeTok{mid\_year =}\NormalTok{ From }\SpecialCharTok{+}\NormalTok{ (To }\SpecialCharTok{{-}}\NormalTok{ From)}\SpecialCharTok{/}\DecValTok{2}\NormalTok{)}

\NormalTok{Forest\_model\_all }\OtherTok{\textless{}{-}} \FunctionTok{gam}\NormalTok{(DeltaQf\_perc }\SpecialCharTok{\textasciitilde{}}\NormalTok{ DeltaF\_perc }\SpecialCharTok{+} 
                    \FunctionTok{s}\NormalTok{(}\FunctionTok{log10}\NormalTok{(Area\_km2), }\AttributeTok{k =} \DecValTok{5}\NormalTok{, }\AttributeTok{bs=}\StringTok{"ts"}\NormalTok{) }\SpecialCharTok{+} 
                    \FunctionTok{s}\NormalTok{(Dryness, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{bs=}\StringTok{"ts"}\NormalTok{ ) }\SpecialCharTok{+} 
                     \FunctionTok{s}\NormalTok{(length, }\AttributeTok{k =} \DecValTok{35}\NormalTok{, }\AttributeTok{bs=}\StringTok{"ts"}\NormalTok{) }\SpecialCharTok{+}
\NormalTok{                    Precip\_data\_type }\SpecialCharTok{+}\NormalTok{  Assessment\_technique }\SpecialCharTok{+}
\NormalTok{                    Forest\_type }\SpecialCharTok{+}
\NormalTok{                    Hydrological\_regime}
\NormalTok{                    , }\AttributeTok{data =}\NormalTok{ All\_data2)}
\CommentTok{\#summary(Forest\_model\_all)}
\CommentTok{\#gam.check(Forest\_model\_all)}
\CommentTok{\#plot(model6\_all)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4231}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1410}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1667}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1282}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1410}}@{}}
\caption{(\#tab:m\_all-linear) Statistical summary for the linear terms the full model}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endhead
\textbf{(Intercept)} & -4.81 & 16.31 & -0.29 & 0.77 \\
\textbf{DeltaF\_perc} & -0.6 & 0.06 & -10.71 & 0 \\
\textbf{Precip\_data\_typeOB} & -21.4 & 13.23 & -1.62 & 0.11 \\
\textbf{Precip\_data\_typeSG} & 9.36 & 15.17 & 0.62 & 0.54 \\
\textbf{Assessment\_techniqueEA, HM} & 20.64 & 42.73 & 0.48 & 0.63 \\
\textbf{Assessment\_techniqueHM} & 22.81 & 11.71 & 1.95 & 0.05 \\
\textbf{Assessment\_techniquePWE} & 30.63 & 11.94 & 2.57 & 0.01 \\
\textbf{Assessment\_techniquePWE,
HM} & 17.42 & 43.26 & 0.4 & 0.69 \\
\textbf{Assessment\_techniqueQPW} & 39.52 & 20.15 & 1.96 & 0.05 \\
\textbf{Assessment\_techniqueQPW,
EA} & 24.39 & 24.42 & 1 & 0.32 \\
\textbf{Assessment\_techniqueSH} & 45.3 & 11.83 & 3.83 & 0 \\
\textbf{Forest\_typeCF} & -9.45 & 7.6 & -1.24 & 0.21 \\
\textbf{Forest\_typeMF} & -8.05 & 7.56 & -1.06 & 0.29 \\
\textbf{Hydrological\_regimeSD} & 3.57 & 9.16 & 0.39 & 0.7 \\
\bottomrule()
\end{longtable}

This is the baseline result as it echoes \citet{zhang2017} but in this case the relationship is tested together with other variables as is discussed in the main paper.

For this model, the focus can be on the linear variables. The variable \texttt{DeltaF\_perc} clearly explains the variation in the change in flow with a p-value of 0. The overall model explains 49 \% of the variation in the data. For comparison across models with different structures and different variables, the Aikaike Information Criteria (AIC) can be used. The AIC of the model is 3181.7600577.

\hypertarget{the-change-in-forest-cover-converted-to-positive-values}{%
\subsection{The change in forest cover converted to positive values}\label{the-change-in-forest-cover-converted-to-positive-values}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Forest\_model\_all2 }\OtherTok{\textless{}{-}} \FunctionTok{gam}\NormalTok{(DeltaQf\_perc }\SpecialCharTok{\textasciitilde{}}\NormalTok{ DeltaF\_perc\_pos }\SpecialCharTok{+}\NormalTok{ Forest\_Sign }\SpecialCharTok{+} 
                    \FunctionTok{s}\NormalTok{(}\FunctionTok{log10}\NormalTok{(Area\_km2), }\AttributeTok{k =} \DecValTok{5}\NormalTok{, }\AttributeTok{bs=}\StringTok{"ts"}\NormalTok{) }\SpecialCharTok{+} 
                    \FunctionTok{s}\NormalTok{(Dryness, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{bs=}\StringTok{"ts"}\NormalTok{ ) }\SpecialCharTok{+} 
                     \FunctionTok{s}\NormalTok{(length, }\AttributeTok{k =} \DecValTok{35}\NormalTok{, }\AttributeTok{bs=}\StringTok{"ts"}\NormalTok{) }\SpecialCharTok{+}
\NormalTok{                    Precip\_data\_type }\SpecialCharTok{+}\NormalTok{  Assessment\_technique }\SpecialCharTok{+}
\NormalTok{                    Forest\_type }\SpecialCharTok{+}
\NormalTok{                    Hydrological\_regime}
\NormalTok{                    , }\AttributeTok{data =}\NormalTok{ All\_data2)}
\CommentTok{\#summary(Forest\_model\_all2)}
\end{Highlighting}
\end{Shaded}

>>>>

\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Testing Forest Area as a random effect},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{Testing Forest Area as a random effect}
    \author[]{R. Willem Vervoort%
  %
  \fnref{1}}
   \ead{willem.vervoort@sydney.edu.au} 
    \author[]{Eliana Nervi}
   \ead{eliananervif@gmail.com} 
    \author[]{Jimena Alonso}
   \ead{jalonso@fing.edu.uy} 
      \cortext[cor1]{Corresponding author}
    \fntext[1]{Corresponding Author}

@@

This document outlines a test suggested in the first review of the manuscript. The AE of the Journal of Hydrology suggested that we test if including the total forest area as a variable would make sense. Not only the \% change in area is important, but what total area of forest was there in the first place.

This test focusses on the large catchment data (\textless{} 1000 km\textsuperscript{2}) for two reasons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  This was a manageable subset and many of the papers are fairly accessible, in contrast to the small catchments.
\item
  Many of the small catchments have 100\% area covered in forest, so are not useful to identify if total area of forest has an impact on the change in flow.
\end{enumerate}

\hypertarget{set-up-the-packages}{%
\subsection{set up the packages}\label{set-up-the-packages}}

\hypertarget{read-in-the-data}{%
\subsection{read in the data}\label{read-in-the-data}}

Read in the data and massage column names to make columns useable in the model.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Data\_wF }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../../data/LCdatawithFA.xlsx {-} LW\_TotForest\_Area.csv"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Rows: 61 Columns: 25
## -- Column specification --------------------------------------------------------
## Delimiter: ","
## chr (10): Watershed name, Forest type, Hydrological regime, yearFarea, Comme...
## dbl (15): Watershed #, Area(km2), Pa(mm), Farea_km2, Perc_Farea_pre, DeltaF_...
## 
## i Use `spec()` to retrieve the full column specification for this data.
## i Specify the column types or set `show_col_types = FALSE` to quiet this message.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# change some column names}
\FunctionTok{names}\NormalTok{(Data\_wF)[}\DecValTok{3}\SpecialCharTok{:}\DecValTok{4}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Area\_km2"}\NormalTok{, }\StringTok{"Pa\_mm"}\NormalTok{)}
\FunctionTok{names}\NormalTok{(Data\_wF)[}\DecValTok{5}\SpecialCharTok{:}\DecValTok{6}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Forest\_type"}\NormalTok{, }\StringTok{"Hydrological\_regime"}\NormalTok{)}
\FunctionTok{names}\NormalTok{(Data\_wF)[}\DecValTok{13}\SpecialCharTok{:}\DecValTok{14}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Precip\_data\_type"}\NormalTok{, }\StringTok{"Assessment\_technique"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{some-changes-to-the-overall-data}{%
\subsection{Some changes to the overall data}\label{some-changes-to-the-overall-data}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  calculating the dryness
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# calculate dryness index}
\NormalTok{Data\_wF }\OtherTok{\textless{}{-}}\NormalTok{ Data\_wF }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Dryness =}\NormalTok{ E0}\SpecialCharTok{/}\NormalTok{Pa\_mm)}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  remove watershed 1 (the Amazon) from the analysis
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Data\_wF }\OtherTok{\textless{}{-}}\NormalTok{ Data\_wF }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Watershed \#}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  include length as a variable
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{All\_data }\OtherTok{\textless{}{-}}\NormalTok{ Data\_wF }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{length =}\NormalTok{ To }\SpecialCharTok{{-}}\NormalTok{ From,}
         \AttributeTok{mid\_year =}\NormalTok{ From }\SpecialCharTok{+}\NormalTok{ (To }\SpecialCharTok{{-}}\NormalTok{ From)}\SpecialCharTok{/}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

Run the model simply as a GLM, transform Area using log10, as in the manuscript.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Forest\_model\_all }\OtherTok{\textless{}{-}} \FunctionTok{gam}\NormalTok{(DeltaQf\_perc }\SpecialCharTok{\textasciitilde{}}\NormalTok{ DeltaF\_perc }\SpecialCharTok{+} 
                    \FunctionTok{log10}\NormalTok{(Area\_km2) }\SpecialCharTok{+} 
\NormalTok{                    Dryness }\SpecialCharTok{+} 
\NormalTok{                    Perc\_Farea\_pre }\SpecialCharTok{+}
\NormalTok{                    Precip\_data\_type }\SpecialCharTok{+}\NormalTok{  Assessment\_technique }\SpecialCharTok{+}
\NormalTok{                    Forest\_type }\SpecialCharTok{+}
\NormalTok{                    Hydrological\_regime}
\NormalTok{                    , }\AttributeTok{data =}\NormalTok{ All\_data)}
\FunctionTok{summary}\NormalTok{(Forest\_model\_all)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## DeltaQf_perc ~ DeltaF_perc + log10(Area_km2) + Dryness + Perc_Farea_pre + 
##     Precip_data_type + Assessment_technique + Forest_type + Hydrological_regime
## 
## Parametric coefficients:
##                             Estimate Std. Error t value Pr(>|t|)   
## (Intercept)                -16.04690   15.08196  -1.064  0.29676   
## DeltaF_perc                 -0.37440    0.11561  -3.239  0.00318 **
## log10(Area_km2)              1.55899    2.31470   0.674  0.50634   
## Dryness                      4.40194    3.21852   1.368  0.18269   
## Perc_Farea_pre               0.03722    0.07544   0.493  0.62572   
## Precip_data_typeOB         -19.28048    6.55235  -2.943  0.00661 **
## Precip_data_typeSG         -12.00304    6.75886  -1.776  0.08702 . 
## Assessment_techniqueEA, HM   1.36550    9.86348   0.138  0.89092   
## Assessment_techniqueHM      15.44379    6.16998   2.503  0.01866 * 
## Assessment_techniqueQPW      2.79556    9.90330   0.282  0.77988   
## Assessment_techniqueSH      18.51105    7.64697   2.421  0.02249 * 
## Forest_typeCF               -9.46812    6.16583  -1.536  0.13628   
## Forest_typeMF                0.27749    4.64399   0.060  0.95279   
## Hydrological_regimeSD       20.97769    6.66044   3.150  0.00397 **
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## 
## R-sq.(adj) =  0.735   Deviance explained = 82.1%
## GCV = 128.37  Scale est. = 84.534    n = 41
\end{verbatim}

This suggest that forest area is not at all significant in the model. In fact, the hydrological modelling technique is much more important. Overall the model explains roughly 82\% of the variation.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gam.check}\NormalTok{(Forest\_model\_all)}
\end{Highlighting}
\end{Shaded}

\includegraphics{Test_model_totalForestArea_files/figure-latex/check-model-percent-1.pdf}

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Model required no smoothing parameter selectionModel rank =  14 / 14
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#plot(Forest\_model\_all)}
\end{Highlighting}
\end{Shaded}

The fit of the model is also well-behaved, as would be expected. The data are mostly from hydrological modelling, stabilising the response of the change in forest to change in streamflow.

\hypertarget{run-the-model-on-forest-area-rather-than}{%
\subsection{run the model on forest area rather than \%}\label{run-the-model-on-forest-area-rather-than}}

Possibly, using the total \% forest is not taking into account the differences in total area of forest. So we can also test with total ha. Note that this can interact with total area: larger catchments will have larger total areas of forest.

We appear to have less data here as \% is reported more, but we can calculate total ha by multiplying \% by total area.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{All\_data }\OtherTok{\textless{}{-}}\NormalTok{ Data\_wF }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Farea\_km2 =} \FunctionTok{ifelse}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(Farea\_km2)}\SpecialCharTok{==}\NormalTok{T,Area\_km2}\SpecialCharTok{*}\NormalTok{Perc\_Farea\_pre}\SpecialCharTok{/}\DecValTok{100}\NormalTok{,Farea\_km2))}

\NormalTok{Forest\_model\_area }\OtherTok{\textless{}{-}} \FunctionTok{gam}\NormalTok{(DeltaQf\_perc }\SpecialCharTok{\textasciitilde{}}\NormalTok{ DeltaF\_perc }\SpecialCharTok{+} 
                    \FunctionTok{log10}\NormalTok{(Area\_km2) }\SpecialCharTok{+} 
\NormalTok{                    Dryness }\SpecialCharTok{+} 
\NormalTok{                    Farea\_km2 }\SpecialCharTok{+}
\NormalTok{                    Precip\_data\_type }\SpecialCharTok{+}\NormalTok{  Assessment\_technique }\SpecialCharTok{+}
\NormalTok{                    Forest\_type }\SpecialCharTok{+}
\NormalTok{                    Hydrological\_regime}
\NormalTok{                    , }\AttributeTok{data =}\NormalTok{ All\_data)}
\FunctionTok{summary}\NormalTok{(Forest\_model\_area)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## DeltaQf_perc ~ DeltaF_perc + log10(Area_km2) + Dryness + Farea_km2 + 
##     Precip_data_type + Assessment_technique + Forest_type + Hydrological_regime
## 
## Parametric coefficients:
##                              Estimate Std. Error t value Pr(>|t|)    
## (Intercept)                -3.021e+01  1.641e+01  -1.841 0.076621 .  
## DeltaF_perc                -3.718e-01  9.294e-02  -4.000 0.000442 ***
## log10(Area_km2)             6.084e+00  3.346e+00   1.818 0.080140 .  
## Dryness                     3.255e+00  3.156e+00   1.031 0.311481    
## Farea_km2                  -6.197e-05  3.922e-05  -1.580 0.125688    
## Precip_data_typeOB         -1.740e+01  6.224e+00  -2.795 0.009431 ** 
## Precip_data_typeSG         -8.667e+00  6.680e+00  -1.297 0.205467    
## Assessment_techniqueEA, HM  4.465e-01  9.411e+00   0.047 0.962513    
## Assessment_techniqueHM      1.436e+01  5.973e+00   2.404 0.023345 *  
## Assessment_techniqueQPW    -6.118e-01  9.487e+00  -0.064 0.949055    
## Assessment_techniqueSH      1.741e+01  7.243e+00   2.404 0.023362 *  
## Forest_typeCF              -8.111e+00  5.993e+00  -1.353 0.187195    
## Forest_typeMF              -4.063e-01  4.433e+00  -0.092 0.927641    
## Hydrological_regimeSD       2.346e+01  6.605e+00   3.552 0.001428 ** 
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## 
## R-sq.(adj) =  0.756   Deviance explained = 83.5%
## GCV = 118.56  Scale est. = 78.075    n = 41
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gam.check}\NormalTok{(Forest\_model\_area)}
\end{Highlighting}
\end{Shaded}

\includegraphics{Test_model_totalForestArea_files/figure-latex/check_resid_area-1.pdf}

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Model required no smoothing parameter selectionModel rank =  14 / 14
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#plot(Forest\_model\_all)}
\end{Highlighting}
\end{Shaded}

Again, not significant and the model is well-behaved. Note that the Hydrological modelling assessment technique (HM) is again highly significant.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

There is no effect of forest area in the relationship between change in flow and percent change in forest cover.

\bibliography{forestandwater.bib}
